Improving credibility and efficiency of decision-making within democratic communities via a statistical quorum

### 2018-04-19 statistical-quorum.tex
theory updates, plus figures

% Document starts
\begin{document}
% Title portion. Note the short title for running heads
\title[Improving credibility and efficiency of decision-making]{Improving credibility and efficiency of decision-making \\ within democratic communities via statistical quorum}

\author{Mitar Milutinovic}
\affiliation{%
  \institution{UC Berkeley}}
\email{mitar@cs.berkeley.edu}
\author{Maxinder S. Kanwal}
\affiliation{%
  \institution{UC Berkeley}}
\email{maxowaxo2013@gmail.com}
\author{David Marn}
\affiliation{%
  \institution{UC Berkeley}}
\email{david.marn@gmail.com}

\begin{abstract}
The Achilles heel of many major democracies is the eternal need for sufficient engagement and attention from the community.
While there have been major efforts towards increasing these attributes within the community, we explore a fundamentally
different approach: For a fixed level of community attention, how can we optimize the allocation of this attention to
maximize productivity?
We propose a novel \emph{statistical quorum}, derived using first principles, as one metric the community can use to
focus attention, in turn making the entire decision-making process dynamic.
We developed a web application centered around this metric and tested it in a 140-person community that holds weekly
councils over the course of 16 weeks.
We found that the community was able to use the web application to simultaneously improve the credibility and efficiency
of their decision-making process.
\end{abstract}

@@

@@

\begin{table}
  \centering
  \begin{small}
  \begin{tabular}{l r r r r r r r}
    {\small\textit{Vote}} & \multicolumn{2}{r}{\small \textit{First option}} & \multicolumn{2}{r}{\small \textit{Second option}} & {\small \textit{Total eligible}} & {\small \textit{Turnout}} & {\small \textit{Decision}} \\
    & & & & & {\small \textit{voters}} & & {\small \textit{credibility}} \\
    \midrule
    Brexit & 17,410,742 & $51.9\%$ & 16,141,241 & $48.1\%$ & 46,500,001 & $72.2\%$ & $59.8\%$ \\
    Presidential elections & 62,984,825 & $48.9\%$ & 65,853,516 & $51.1\%$ & 230,585,915 & $55.9\%$ & $52.6\%$ \\
  \end{tabular}
  \end{small}
  \bigskip\caption{Votes cast in Brexit, a 2016 referendum for United Kingdom's withdrawal from the European Union~\protect\cite{brexitresults},
  and 2016 USA presidential elections, popular vote~\protect\cite{usaresults1,usaresults2}. Simplified to only two options and without
  invalid votes. The last column shows computed decision credibility as described in Section~\ref{sec:decision-credibility}. We see that
  credibility is low, which we can mostly attribute to the fact that there is small difference in votes between both options so
  with more votes any option could end up winning.}\label{tab:example}
  \vspace{-0.3cm}
\end{table}

@@

Our contributions are:

\begin{itemize}
\item We developed a framework of computing statistical credibility of results
based on sample size, votes cast, and voting duration.
\item Based on this framework we developed a novel concept of \emph{statistical quorum}
as a threshold of votes required to deem results credible and legitimate.
\item We implemented web application \emph{PeerMind},
integrating credibility of results and statistical quorum, and designed
a workflow of its use in a decision making processes.
\item We deployed the application in a democratically run community of 140
members for 16 weeks.
\item Using the analysis of decision making process before and after introduction
of the application, using surveys, and using semi-structured interviews, we show
how use of statistical quorum increases credibility of decisions,
improves utilization of collective attention, increases and diversifies overall participation,
and increases number of decisions made.
\end{itemize}

@@

%\todo[inline]{Threshold and quorum at \url{https://www.publications.parliament.uk/pa/ld200910/ldselect/ldconst/99/99.pdf}.
%How do some claim that quorum are in fact bad, and counter democratic, but do not offer a solution how to evaluate decisions.}
%
%\todo[inline]{Voting auditing work.}
%
%\todo[inline]{Delegation as way to increase legitimacy, Liquid Democracy, etc.}

\section{Decision Credibility and Statistical Quorum}
\label{sec:decision-credibility}

%Range voting (also known as score voting) is voting where members of a community express their positions on open proposals by assigning
%each a score on a fixed scale, in the context of this paper, between $-1$ and $+1$.
%A proposal passes if the average score of all votes on a position is positive.
%When multiple competing proposals are being voted on, the one with highest average score wins.
%
%Prior work has shown that this approach generalizes many previous approaches to voting and allows members to express their
%positions in a highly informative and precise manner.
%We find it especially suitable for a web application because many web users are used to use it when ranking products
%and giving other ``five-star'' feedback.
%Moreover, if later on a new proposal is added to a set of existing proposals on which voting was already open, only the
%new proposal needs to be evaluated by members, as opposed to alternative voting systems where the relative rankings of
%each proposal would need reassessing (assuming members use a consistent personal scale when voting).

A statistical quorum uses statistical hypothesis testing to infer the credibility of the outcome
of an ongoing voting before the community adopts the outcome as being truly representative of all
members' positions.
For example, in a community of 100 members, if 35 members support a specific proposal and 0 oppose,
then we can conclude with high probability that, had all 100 members taken the time to vote, a
majority of them would be in support of the proposal.
Hence, despite the reality of 65 members not having voted, the outcome of the 35-member vote is
credible enough to be taken as representative of the entire community's desire.
On the other hand, if 80 members vote, where 41 support the proposal and 39 oppose, then we
cannot yet conclude that the decision to accept the proposal is credible enough for the entire
community to officially adopt, despite the significant participation in voting.
Instead, the suggested course of action would be to continue gathering more votes and/or further
discussing the proposal until a credible outcome is established.

@@

The primary motivation behind using a statistical quorum is to guarantee a level of credibility
before making a decision on behalf of the entire community.

@@

\subsection{Assumptions}
We assume that there is a fixed latent bias for the support of a proposal within the community.
We take each individual vote to be independent and identically distributed (i.i.d.) according
to a Bernoulli distribution around this bias.
It should be noted that the independence assumption may not hold in some settings, such as when
votes are accumulated over time (one may observe surges of correlated votes).
We will partially address this when refining our original formulation of a statistical quorum
into a \textit{dynamic} statistical quorum.

Without having collected any votes, we assume that the community bias may be any value between
$0\%$ and $100\%$ with equal probability (uniform prior).

\subsection{Voting}
We allow for range voting where each member of the community may choose any real-valued vote
$v \in [-1,+1]$ to represent their vote, with $v = -1$ representing full opposition and $v=+1$
representing full support.
In most applications, it is sensible to quantize $v$ into a fixed number of bins spanning
the full range, hence our treatment of $v$ as a discrete random variable.
Because we treat individual votes as a Bernoulli random variable, one can consider $v$ as
an individual's average vote.
In the following subsection, we rescale $v$ to be in the range $[0,1]$ via the
notation $\frac{v + 1}{2}$.

@@

After collecting an initial set of votes from members, we want to use that data to construct
a prior distribution to help estimate the posterior probability that the proposal is acceptable
to the entire community.
Because the sum of the collected votes follows a binomial distribution, a commonplace and
convenient choice for the form of the prior on the community's bias is the Beta distribution
(due to its conjugacy with the binomial distribution).
Together, they form the Beta-binomial distribution (related to the P\'olya urn model).

@@

Let
\begin{description}
\item[$C = $] $\{X_1,\ldots, X_n\}$, the set of competing proposals,
\item[$N = $] population size,
\item[$X_i = $] $\{v_1, \ldots, v_{m_i}\}$, the $m_i$ raw votes for proposal $i$, after excluding
abstentions,
\item[$N_i = $] $(N - \textrm{number of abstentions on } X_i)$, the effective population voting
on proposal $X_i$,
\item[$\alpha_i = $] $\sum\limits_{v \in X_i}\left(\frac{v+1}{2}\right)$, the effective number
of ``yes'' votes on proposal $X_i$,
\item[$\beta_i = $] $m_i - \alpha_i$, the effective number of ``no'' votes on proposal $X_i$,
\item[$K_i = $] $N_i - m_i$, the number of remaining votes yet to be cast on $X_i$,
\item[$p_i = $] true latent community bias towards voting ``yes'' on $X_i$.
\end{description}

In Eqs.~\eqref{EQ2}, \eqref{EQ3}, \eqref{EQ5}, \eqref{EQ6}, \eqref{EQ7} that follow, we refer the
reader to the \hyperref[sec:derivation]{Appendix} for complete derivation details and justifications.

When considering a single proposal $X_i$ in isolation, we wish to compute:
\begin{align}\label{EQ1}
\Pr(\alpha^*_i > T_i \mid X_i),
\end{align}
where
\begin{description}
\item[$\alpha^*_i = $] the total number of ``yes'' votes for $X_i$ if we were to sample the entire
population,
\item[$T_i = $] $f \cdot N_i$, for some fraction $f$ corresponding to the type of majority needed
(e.g., $f=\frac{1}{2}$ for a simple majority).
\end{description}
Using a desirable prior
\begin{align}\label{EQ2}
p_i \mid X_i \sim \operatorname{Beta}(1+\frac{\alpha_i}{K_i},1+\frac{\beta_i}{K_i}),
\end{align}
corresponding to a uniform prior when no votes have been cast ($\alpha_i = \beta_i = 0$), we can
express \eqref{EQ1} as:
\begin{align}\label{EQ3}
\Pr(\alpha^*_i > T_i \mid X_i) &= 1 - {CDF}_{\nu_i}(V_i)
\end{align}
for
\begin{align*}
&\nu_i \sim \operatorname{Beta-Binomial}(1 + \frac{\alpha_i}{K_i}c, 1 + \frac{\beta_i}{K_i}c, K_i),\\
&{CDF}_{\nu_i}(V_i)  = \Pr(\nu_i \leq V_i),
\end{align*}
\todo[inline]{Note that we choose $c=10$ because it leads to intuitive results.}
where
\begin{description}
\item[$V_i = $] $\lfloor{T_i - \alpha_i}\rfloor$, the maximum number of additional ``yes''
votes $X_i$ could receive such that it would remain below threshold.
\end{description}

When considering a proposal $X_i$ relative to other proposals in $C$, we wish to compute
\begin{align}\label{EQ4}
\Pr\left(\bigwedge_{j \neq i} (\alpha^*_i > \alpha^*_j) \mid C\right)
\end{align}

Under the assumption that the community's votes on proposals are independent across proposals
(which, in many instances, will actually give a lower bound on \eqref{EQ3} when a community's
support of one proposal is negatively correlated with its support on a competing proposal),
we can decompose \eqref{EQ3} as follows:
\begin{align}\label{EQ5}
& \Pr\left(\bigwedge_{j \neq i} (\alpha^*_i > \alpha^*_j) \mid C\right) \nonumber \\
= & \prod_{j \neq i}\left(1 - \sum_{k=\lceil\delta_{ij}\rceil}^{K_j} \Pr(\nu_j = k) \cdot {CDF}_{\nu_i}(\alpha_j + k) \right)
\end{align}
where
\begin{description}
\item[$\delta_{ij} = $] $\alpha_i - \alpha_j$, the number of effective ``yes'' votes $X_i$
leads $X_j$ by.
\end{description}

Using Eqs.~\eqref{EQ2} and \eqref{EQ4}, we can numerically compute the credibility of a
vote's outcome in two key scenarios relevant for democratic communities.
For example, a community may only consider accepting a proposal when a vote's credibility
is above $90\%$.

\subsection{Dynamic Statistical Quorum}
Oftentimes not all proposals are relevant to the entire community.
Instead, a niche proposal may only affect a subset of the community, and hence to require much
of the community to indicate their indifference and/or decide their opinion is a waste of their
attentional budget.
While the underlying formulation for a statistical quorum was to establish credibility guarantees
before making a decision, a natural consequence in common scenarios of the kind just described
leads to a reduction in decision-making efficiency.
How can we maintain credibility guarantees while simultaneously boosting efficiency?

The solution is to infer the \textit{relevant} population size that cares about a given proposal
so that we may establish credibility within only the subset population (naturally requiring less
global attention).
By replacing the absolute upper bound on this relevant population size ($N_i$) with a high
probability upper bound, we can continue to use the same formula presented in Eqs.~\eqref{EQ2}
and \eqref{EQ4}, but with more relevant parameters (namely, a smaller $N_i$, which in turn
modifies $T_i$, $K_i$, and $V_i$).

@@

For our purposes, we make a natural choice,
\begin{align}\label{EQ6}
\lambda (t) = \lambda_0 {(\gamma)}^{\frac{t}{K_i/N_i}},
\end{align}
where
\begin{description}
\item[$\lambda_0 =$] a time-averaged rate of vote arrivals (preferably over a unit of time that
sufficiently smooths the historical rate),
\item[$\gamma =$] a decay factor ($0 < \gamma < 1$) ideally tuned to fit real historical data
for the community via a regression analysis.
\end{description}
Now, for example, to upper bound with $95\%$ probability the remaining number of votes still
pending, we wish to solve for $\hat{K_i}$ such that the following inequality is satisfied:
\begin{align}\label{EQ7}
\lim_{t\rightarrow \infty}\Pr\left(N(t)\leq \hat{K_i}\right) & = \sum_{n=0}^{\hat{K_i}} \frac{\Lambda^n}{n!}e^{-\Lambda} \geq 95\%,
\end{align}
where $\Lambda = \frac{\lambda_0 \cdot K_i/N_i)}{\ln{1/\gamma}}$.
Upon solving for $\hat{K_i}$, we update $N_i \leftarrow \min\{N_i,\alpha_i + \beta_i + \hat{K_i}\}$,
and can proceed to check for a statistical quorum using Eq.~\ref{EQ5}.

@@

For the baseline, we extracted similar information from council minutes: how long did the councils last,
how many members participated, how many proposals were proposed and how many passed, and vote tallies for them.

@@

We built a full-fletched mobile-friendly web application using Meteor web framework~\cite{meteor}.
Meteor framework supports reactivity which makes the application update user interface based on actions
of other users in real time. This helps users see comments by other users as they happen, but can also see
progress during voting in real time.

@@

An important design decision in the scope of this paper is that during voting, users can see the current
results of voting in real time. Results are shown on a chart depicting average score through time
as seen in Figure~\ref{fig:results}. Users can see both the current average score of all votes (score
larger than 0 means that a proposal is passing) and the credibility of current result. A community can decide
how high credibility it requires to deem the result passing statistical quorum. During our evaluation
we used 0.90 as statistical quorum.

The rationale for displaying the current results during voting is to give users feedback how voting
is progressing and allow them to engage more with the discussion. Moreover, displaying the results
also serves as additional safeguard against potential manipulations of statistical quorum through malicious
coordination.

After voting passes the statistical quorum, a moderator can close voting and a discussion item.

@@

PeerMind source code is available under open source licence at \url{https://redacted.for.blind.review} and
anonymized dataset of interactions with the app, including votes, is available at \url{https://redacted.for.blind.review}.

>>>>

% Document starts
\begin{document}
% Title portion. Note the short title for running heads
\title{Improving credibility and efficiency of decision-making within democratic communities via a statistical quorum}

\author{Mitar Milutinovic}
\affiliation{%
  \institution{UC Berkeley}}
\email{mitar@cs.berkeley.edu}
\author{Maxinder S. Kanwal}
\affiliation{%
  \institution{UC Berkeley}}
\email{mkanwal@berkeley.edu}
\author{David Marn}
\affiliation{%
  \institution{UC Berkeley}}
\email{david.marn@gmail.com}

\begin{abstract}
The Achilles heel of many major democracies is the perpetual need for sufficient engagement and attention from the community. While there have been major efforts towards increasing these attributes within the community, we explore a fundamentally different approach: For a fixed level of community attention, how can we optimize the allocation of this attention to maximize productivity? We propose a novel \textit{statistical quorum}, derived using first principles, as one metric the community can use to focus attention, in turn making the entire decision-making process dynamic. We developed a web application centered around this metric and tested it in a 140-person community that holds weekly councils over the course of 16 weeks. We found that the community was able to use the web application to simultaneously improve the credibility and efficiency of their decision-making process.
\end{abstract}

@@

Two of the most shocking democratic votes in recent history occurred in 2016: The referendum vote for Britain to leave the European Union (“Brexit”) and the election of Donald Trump as President of the United States. Both votes were decided by the slimmest of margins, but more significantly, both votes also had poor turnout. The Achilles heel of many major democracies is the eternal need for sufficient engagement and attention from the community. Hence, there have been major efforts towards increasing these attributes within the community. We believe this is a fallacy. Democracy need not suffer under low engagement — in fact, our goal should be to make democracy work with minimal engagement. The ideal governance system is one where citizens do not have to be constantly involved, but the system is still acting as the citizens would like it to.
We explore a fundamentally different approach to making democracy work: \textit{For a fixed level of community attention, how can we optimize the allocation of this attention to maximize productivity?} Can we increase the credibility of decisions made within a democracy while simultaneously skirting the need for an increase in attention? We formulated a novel ‘statistical quorum’, derived using first principles, as one metric the community can use to focus attention, in turn making the entire decision-making process dynamic. We have developed a web application centered around this metric and tested it in a 140-person community that holds weekly councils over the course of 16 weeks. We found that the community was able to use the web application to simultaneously improve the legitimacy and efficiency of their decision-making process.
Our prototyped web application for collective decision-making contrasts with existing similar solutions. Our web application integrates new ideas in how to improve this process: use of \textbf{range voting}, a \textbf{dynamic statistical quorum}, and \textbf{voter delegation}. The broad idea is that, instead of just trying to move the voting process online — something many others have done previously — we in fact wish to add value by voting online. Specifically, we can compute results faster, more accurately, and securely, we can use powerful statistical methods to estimate important quantities for the entire voting population, we can allow for members to entrust their vote to others who have more expertise on specific topics than they themselves do, and we can promote clarity of thought and complete understanding among all members for all voting items.
\textit{Range voting} is voting where members of a community express their positions on open proposals by assigning each a score between –1 and +1. Prior work has shown that this approach generalizes many previous approaches to voting and allows members to express their positions in a highly informative and precise manner. Moreover, if a new proposal is added to a set of existing proposals on which voting was already open, only the new proposal needs to be evaluated by members, as opposed to alternative voting systems where the relative rankings of each proposal would need reassessing.
A \textit{dynamic statistical quorum} uses statistical hypothesis testing to compute the minimum number of required votes needed to achieve a desired level of confidence before a community adopts an outcome as being truly representative of its members’ positions. For example, in a community of 100 members, if 20 members all vote in full support of a proposal, then we can conclude with very high confidence that, had all 100 members actually voted, a majority of them would support the proposal (despite in actuality 80 members having not yet voted). On the other hand, if 80 members vote with 41 voting in favor of the proposal and 39 voting in opposition, then the community cannot yet conclude anything with high confidence (despite the high voter turnout). Hence, the suggested course of action would be for the community to gather more votes and continue discussing the proposal. By utilizing a dynamic statistical quorum, we can direct attention and participation within the community towards contentious proposals, while reducing wasted attention directed towards proposals that already have overwhelming support.
\textit{Voter delegation} addresses two questionable assumptions about current democratic systems: (1) participants have the time to participate and (2) participants have the knowledge to participate. Not all members of a community can participate at all times and not all members have the time to educate themselves about each and every decision being made. By allowing members to define delegates to whom they can give their own vote to — likely those who share their values — their proxy vote can be inferred from their delegates’ votes. In this way, their values are still voiced even if they do not vote directly. If at any point a member decides to vote themselves, then their direct vote is used instead.
All of the above features are combined into a guided process to help community members participate in decision-making. The process starts with a new discussion item where members can comment and discuss a proposal. Moderators then summarize the discussion into three types of points: “in favor,” “against,” and “other”. Members can then use those summaries to help form an opinion before they vote, without having to read every discussion comment made in full detail, thus lowering the barrier of entry into any discussion. As members vote, the app shows the community the average score and confidence in real-time. When the confidence passes a preset threshold, the proposal can be firmly accepted by the community.

@@

\begin{table}
  \centering
  \begin{small}
  \begin{tabular}{l r r r r r r r}
    {\small\textit{Vote}} & \multicolumn{2}{r}{\small \textit{First option}} & \multicolumn{2}{r}{\small \textit{Second option}} & {\small \textit{Total eligible}} & {\small \textit{Turnout}} & {\small \textit{Decision}} \\
    & & & & & {\small \textit{voters}} & & {\small \textit{credibility}} \\
    \midrule
    Brexit & 17,410,742 & $51.9\%$ & 16,141,241 & $48.1\%$ & 46,500,001 & $72.2\%$ & $59.8\%$ \\
    Presidential elections & 62,984,825 & $48.9\%$ & 65,853,516 & $51.1\%$ & 230,585,915 & $55.9\%$ & $52.6\%$ \\
  \end{tabular}
  \end{small}
  \bigskip\caption{Votes cast in Brexit, a 2016 referendum for United Kingdom's withdrawal from the European Union~\protect\cite{brexitresults},
  and 2016 USA presidential elections, popular vote~\protect\cite{usaresults1,usaresults2}. Simplified to only two options and without
  invalid votes. The last column shows computed decision credibility as described in Section~\ref{sec:decision-credibility}. We see that
  credibility is low, which we can mostly attribute to the fact that there is small difference in votes between both options so with a larger turnout, any option could have ended up winning.}\label{tab:example}
  \vspace{-0.3cm}
\end{table}

@@

Our main contributions are:

\begin{enumerate}
\item We developed a framework of computing statistical credibility of results
based on sample size, votes cast, and voting duration.
\item Based on this framework we developed the novel concept of a \emph{statistical quorum} as a threshold of votes required to deem results credible and legitimate.
\item We implement a web application \emph{PeerMind},
integrating credibility of results and statistical quorum, and designed
a workflow of its use in decision making processes.
\item We deployed the application in a democratically run community of 140
members for 16 weeks.
\item Using the analysis of decision making process before and after introduction
of the application, using surveys, and using semi-structured interviews, we show
how use of statistical quorum increases credibility of decisions,
improves utilization of collective attention, increases and diversifies overall participation,
and increases number of decisions made.
\end{enumerate}

@@

%\todo[inline]{Threshold and quorum at \url{https://www.publications.parliament.uk/pa/ld200910/ldselect/ldconst/99/99.pdf}.
%How do some claim that quorum are in fact bad, and counter democratic, but do not offer a solution how to evaluate decisions.}
%
%\todo[inline]{Voting auditing work.}
%
%\todo[inline]{Delegation as way to increase legitimacy, Liquid Democracy, etc.}
\section{Related Work}
\label{sec:related-work}

%\todo[inline]{Repeat references from "There have been many improvements proposed to this basic form and various analysis made."?}
%
%\todo[inline]{Threshold and quorum at \url{https://www.publications.parliament.uk/pa/ld200910/ldselect/ldconst/99/99.pdf}.
%How do some claim that quorum are in fact bad, and counter democratic, but do not offer a solution how to evaluate decisions.}
%
%\todo[inline]{Voting auditing work.}
%
%\todo[inline]{Delegation as way to increase legitimacy, Liquid Democracy, etc.}
%
%\section{PeerMind Web Application}
%\label{sec:peermind}
%
%\todo[inline]{Something we need to think about is the level of detail we want to go into regarding the implementation/features of PeerMind beyond statistical quorum and range voting.  I think we might want to focus on only those two features for this paper.\\
%I think PeerMind as a comprehensive platform should have its own paper where all of the data is analyzed across the board.  This would be a much more typical HCI paper.  (If I can become president of Ridge in a future semester, we'll have another chance to run a clean, well-designed experiment with a much more developed version of PeerMind.)\\
%The message in the current paper might be clearest if we keep the scope focused on the theory + evaluation/verification.  We should make sure we're not trying to pack too much into one paper.
%\\---Max}
%
%We built a full-fletched real-time web application for collective decision making and voting using Meteor
%framework~\cite{meteor}.
%Users can initiate new discussions which leads them to the main user interface for each discussion as seen
%in Figure~\ref{fig:peermind}.
%
%\begin{figure*}
%  \centering
%  \includegraphics[width=1.75\columnwidth]{figures/peermind.png}
%  \caption{The main PeerMind interface of a discussion. There is initial discussion description on the top and under it there
%  are three columns. A commenting sections for users to discuss it and proposed decisions, a second column where moderators can
%  summarize arguments made during discussion, and a list of proposals to address the discussion.
%  Proposals can be in different states: a proposal being discussed, a proposal being voted on, and with voting concluded.}\label{fig:peermind}
%\end{figure*}
%
%%and discuss them with others online in real-time and with rich-text content.
%%They can attach images and files using drag\&drop. Moderators can summarize discussions into three groups
%
%\todo[inline]{Present how users can see ongoing results during voting. Show an image of active ballot, but with expanded view for current results.}
%
%%Range voting (also known as score voting) is voting where members of a community express their positions on open proposals by assigning
%%each a score on a fixed scale, in the context of this paper, between $-1$ and $+1$.
%%A proposal passes if the average score of all votes on a position is positive.
%%When multiple competing proposals are being voted on, the one with highest average score wins.
%%
%%Prior work has shown that this approach generalizes many previous approaches to voting and allows members to express their
%%positions in a highly informative and precise manner.
%%We find it especially suitable for a web application because many web users are used to use it when ranking products
%%and giving other ``five-star'' feedback.
%%Moreover, if later on a new proposal is added to a set of existing proposals on which voting was already open, only the
%%new proposal needs to be evaluated by members, as opposed to alternative voting systems where the relative rankings of
%%each proposal would need reassessing (assuming members use a consistent personal scale when voting).
%
%\todo[inline]{The app has features both for use in online and offline setting. This allows one to compute
%statistical quorum also for an in-person meeting because math is now more complicated than just counting hands.
%We focus on online experience in this paper.}
%
%\todo[inline]{Security is out of the scope of this work. We assume here that users and their identities are known and are managed
%by a community's admin. There are many communities where their decision making is public with public minutes and this is who
%the app is currently targeting.}

\section{Decision Credibility and Statistical Quorum}
\label{sec:decision-credibility}
A statistical quorum uses statistical hypothesis testing to infer the credibility of the outcome of an open vote before the community adopts the outcome as being truly representative all members' positions.
For example, in a community of 100 members, if 35 members support a specific proposal and 0 oppose, then we can conclude with high probability that, had all 100 members taken the time to vote, a majority of them would be in support of the proposal.  Hence, despite the reality of 65 members not having voted, the outcome of the 35-member vote was credible enough to be taken as representative of the entire community's desire.
On the other hand, if 80 members vote, where 41 support the proposal and 39 oppose, then we cannot yet conclude that the decision to accept the proposal is credible enough for the entire community to officially adopt, despite the significant participation in the vote.
Instead, the suggested course of action would be to continue gathering more votes and/or further discussing the proposal until a credible outcome is established.

@@

The primary motivation behind using a statistical quorum is to guarantee a level of credibility before letting a subset of members act on behalf of the entire community.

@@

\subsection{Assumptions}
We assume that there is a fixed latent bias for the support of a proposal within the community.  We take each individual vote to be independent and identically distributed (i.i.d.) according to a Bernoulli distribution around this bias.  It should be noted that the independence assumption may not hold in some settings, such as when votes are accumulated over time (one may observe surges of correlated votes).  We will attempt to address this when refining our original formulation of a statistical quorum into a \textit{dynamic} statistical quorum.

Without having collected any votes, we assume that the community's overall bias may be any value between $0\%$ and $100\%$ with equal probability (uniform prior).

\subsection{Voting}

%\url{https://youtu.be/Q60ZXoXP6Hg}

We allow for range voting where each member of the community may choose any real-valued vote $v \in [-1,+1]$ to
represent their vote, with $v = -1$ representing full opposition and $v=+1$ representing full support.  In most applications, it is sensible to quantize $v$ into a fixed number of bins spanning the full range, hence our treatment of $v$ as a discrete random variable.  Because we treat individual votes as a Bernoulli random variable, one can consider $v$ as an individual's average vote.  In the theory subsection below, we rescale $v$ to be in the range $[0,1]$ via the notation $\frac{v + 1}{2}$.

@@

\todo[inline]{Hyperparameters: Explain not only what you did, but also why you did it, so that readers (including reviewers) can be convinced that you made appropriate choices.}

After collecting an initial set of votes from members, we want to use that data to construct a prior distribution to help estimate the posterior probability that the proposal is acceptable to the entire community.  Because the sum of  the collected votes follows a binomial distribution, a commonplace and convenient choice for the form of the prior on the community's bias is the Beta distribution (due to its conjugacy with the binomial distribution).  Together, they form the Beta-binomial distribution (also known as the P\'olya urn model).

@@

Let

\begin{itemize}
\item[$C = $] $\{X_1,\ldots, X_n\}$, the set of competing proposals,
\item[$N = $] population size,
\item[$X_i = $] $\{v_1, \ldots, v_{m_i}\}$, the $m_i$ raw votes for proposal $i$, after excluding abstentions,
\item[$N_i = $] $(N - \textrm{number of abstentions on } X_i)$, the effective population voting on proposal $X_i$,
\item[$\alpha_i = $] $\sum\limits_{v \in X_i}\left(\frac{v+1}{2}\right)$, the effective number of ``yes'' votes on motion $X_i$,
\item[$\beta_i = $] $m_i - \alpha_i$, the effective number of ``no'' votes on motion $X_i$,
\item[$K_i = $] $N_i - m_i$, the number of remaining votes yet to be cast on $X_i$,
\item[$p_i = $] true latent community bias towards voting ``yes'' on $X_i$.
\end{itemize}

In Eqs. \eqref{eq2}, \eqref{eq3}, \eqref{eq5}, \eqref{eq6}, \eqref{eq7} that follow, we urge the reader to refer to the \hyperref[sec:derivation]{Appendix} for complete derivation details, including motivations and justifications.

When considering a single proposal $X_i$ in isolation, we wish to compute:
\begin{align}\label{eq1}
\Pr(\alpha^*_i > T_i \mid X_i),
\end{align}
where
\begin{itemize}
\item[$\alpha^*_i = $] the total number of ``yes'' votes for $X_i$ if we were to sample the entire population,
\item[$T_i = $] $f \cdot N_i$, for some fraction $f$ corresponding to the type of majority needed for the vote
(e.g., $f=\frac{1}{2}$ for a simple majority).
\end{itemize}
Using a desirable prior 
\begin{align}\label{eq2}
p_i \mid X_i \sim \operatorname{Beta}(1+\frac{\alpha_i}{K_i},1+\frac{\beta_i}{K_i}),
\end{align}
corresponding to a uniform prior when no votes have been cast ($\alpha_i = \beta_i = 0$), we can express \eqref{eq1} as:
\begin{align}\label{eq3}
\Pr(\alpha^*_i > T_i \mid X_i) &= 1 - {CDF}_{\nu_i}(V_i)
\end{align}
for
\begin{align*}
&\nu_i \sim \operatorname{Beta-Binomial}(1 + \frac{\alpha_i}{K_i}c, 1 + \frac{\beta_i}{K_i}c, K_i),\\
&{CDF}_{\nu_i}(V_i)  = \Pr(\nu_i \leq V_i),
\end{align*}
where
\begin{itemize}
\item[$V_i = $] $\lfloor{T_i - \alpha_i}\rfloor$, the maximum number of additional ``yes'' votes $X_i$ could receive such that it would remain below threshold.,
\item[$c = $] $10$, a hyperparameter chosen by the community based on their relative concern about false alarms versus (temporarily) missed detection.
\end{itemize}

When considering a proposal $X_i$ relative to other proposals in $C$, we wish to compute
\begin{align}\label{eq4}
\Pr\left(\bigwedge_{j \neq i} (\alpha^*_i > \alpha^*_j) \mid C\right)
\end{align}

Under the assumption that the community's votes on motions are independent across motions (which, in many instances, will actually give a lower bound on \eqref{eq3} when a community's support of one proposal is negatively correlated with its support on a competing proposal), we can decompose \eqref{eq3} as follows:
\begin{align}\label{eq5}
& \Pr\left(\bigwedge_{j \neq i} (\alpha^*_i > \alpha^*_j) \mid C\right) = \prod_{j \neq i}\left(1 - \sum_{k=\lceil\delta_{ij}\rceil}^{K_j} \Pr(\nu_j = k) \cdot {CDF}_{\nu_i}(k-\delta_{ij}) \right)
\end{align}
where
\begin{itemize}
\item[$\delta_{ij} = $] $\alpha_i - \alpha_j$, the number of effective ``yes'' votes $X_i$ leads $X_j$ by.
\end{itemize}

Using Eqs. \eqref{eq2} and \eqref{eq4}, we can numerically compute the credibility of a vote's outcome in two key scenarios relevant for democratic communities.  For example, a community may only consider accepting a proposal when a vote's credibility is above $90\%$.

\subsection{Dynamic Statistical Quorum}
Oftentimes, not all proposals are relevant to the entire community.  Instead, a niche proposal may only affect a subset of the community, and hence to require much of the community to indicate their indifference and/or decide their opinion is a waste of the their fixed attentional budget.  While the underlying formulation for a statistical quorum was to establish credibility guarantees before taking action, a natural consequence in common scenarios of the kind just described leads to a reduction in decision-making efficiency.  How can we maintain credibility guarantees while simultaneously boosting efficiency?

Our proposed solution is to infer the \textit{relevant} population size that cares about a given proposal so that we may establish credibility within only the subset population (naturally requiring less global attention).  By replacing the absolute upper bound on this relevant population size ($N_i$) with a high probability upper bound, we can continue to use the same formula presented in Eqs. \eqref{eq2} and \eqref{eq4}, but with more relevant parameters (namely, a smaller $N_i$, which in turn modifies $T_i$, $K_i$, and $V_i$).

@@

For our purposes, we make a natural choice,
\begin{align}\label{eq6}
\lambda (t) = \lambda_0 {(\gamma)}^{\frac{t}{K_i/N_i}},
\end{align}
where
\begin{itemize}
\item[$\lambda_0 =$] a time-averaged rate of vote arrivals (preferably over a unit of time that sufficiently smooths the historical rate),
\item[$\gamma =$] a decay factor ($0 < \gamma < 1$) ideally tuned to fit real historical data for the community via a regression analysis.
\end{itemize}
Now, for example, to upper bound with $95\%$ probability the remaining number of votes still pending, we wish to solve for $\hat{K_i}$ such that the following inequality is satisfied:
\begin{align}\label{eq7}
\lim_{t\rightarrow \infty}\Pr\left(N(t)\leq \hat{K_i}\right) & = \sum_{n=0}^{\hat{K_i}} \frac{\Lambda^n}{n!}e^{-\Lambda} \geq 95\%,
\end{align}
where $\Lambda = \frac{\lambda_0 \cdot K_i/N_i}{\ln{(1/\gamma)}}$.
Upon solving for $\hat{K_i}$, we update $N_i \leftarrow \min\{N_i,\alpha_i + \beta_i + \hat{K_i}\}$, and can proceed to check for a statistical quorum using Eq. \ref{eq5}.

@@

For the baseline, we extracted similar information from council minutes: how long did the councils last, how many
discussions were opened, how many comments by which members were made, how many proposals were proposed,
and vote tallies for them.

@@

We built a full-fletched mobile-friendly web applicationusing Meteor web framework~\cite{meteor}.
Meteor framework supports reactivity which makes the application update user interface based on actions
of other users in real time. This helps users see comments by other users as they happen, but can also see
progress during voting in real time.

@@

An important design decision in the scope of this paper is that during voting, users can see the current
results of the vote in real time. Results are shown on a chart depicting average score through time
as seen in Figure~\ref{fig:results}. Users can see both the current average score of all votes (score
larger than 0 means that a proposal is passing) and the credibility of current result. A community can decide
how high credibility it requires to deem the result passing statistical quorum. During our evaluation
we used 0.90 as statistical quorum.

The rationale for displaying the current results during voting is to give users feedback how the vote
is progressing and allow them to engage more with the discussion. Moreover, displaying the results
also serves as additional safeguard against potential manipulations of statistical quorum through malicious
coordination.

After the vote passes the statistical quorum, a moderator can close voting and a discussion item.

@@

PeerMind source code is available under open source licence at \url{http://redacted.for.blind.review} and
anonymized dataset of interactions with the app, including votes, is available at \url{http://redacted.for.blind.review}.

\begin{figure}[ht]
\centering
\includegraphics[width=0.5\textwidth]{figures/peermind.png}
\caption{User interface of PeerMind web application. One can see three column structure, with commenting
section on the left, summary points in the middle, and proposed proposals with voting on the right.}
\label{fig:peer-mind}
\end{figure}
